---
title: "Comapre Flows from Updated WDM files"
author: "Kevin Brannan"
date: 'Wednesday, June 10, 2015'
output: word_document
---

I compared the simulated flows from simulations using the original and updated WDM files for Big Elk Creek. I compared the total volumes for the entire simulation periods, summers, winters, and storms. I used the definitions for periods from the original hydrologic calibration (see table below). The scripts used to do this analysis is avalable at: https://github.com/kbrannan/BigElkCompFlows

Packages used:
```{r}
library(doBy)
library(ggplot2)
library(knitr)
library(xtable)
```

Table 1. Dates and periods for the simulation, summer and winter.

Period                            Begin-End 
--------------------------------  ----------------------------
Original WDM simulation           1995-10-01 to 2010-12-31
Updated WDM simulation            1995-10-01 to 2014-05-31
Summer*                           June through August
Winter*                           September through May

* These were the poriods used in the hydrologic calibration

Set path for analysis and load data from previous runs
```{r}
# load data
chr.run.path <- "C:/Temp/PEST/BigElkPEST/hydrology/UpdatedMetWDM"
load(file=paste0(chr.run.path,"/flowdata.RData"))
```

Create summary table of the runs
```{r}
# convert drainage area of Big Elk Creek from sqr mi to acres-ft using 1 acre = 43560 sqr ft
df.summ00 <- summaryBy((value/43560) ~ file + name, 
                       data=df.data.flow[df.data.flow$type=="V_TABLE" & 
                                           df.data.flow$comments=="begin date",],
                       FUN=c(sum,mean), var.names="vol.ac.ft")
df.summ00 <- cbind(df.summ00,period=factor(gsub("^.*_","", as.character(df.summ00$name)),levels=unique(gsub("^.*_","", as.character(df.summ00$name)))))
```

Here is what the summary looks like:
```{r, results='asis'}
kable(df.summ00,format="markdown",row.names=FALSE)
```

I added a new variable to the summary that is clearer than the "file" variable used in the original data (See Table 2).

Table 2. Cross reference of run name with wdm file and simulation period used.

Run name         WDM file     Simulation Period         Description
---------------  -----------  ------------------------  -----------------------------
Obs              NA           NA                        Observed flow data
Original         Original     1995-10-01 to 2010-12-31  Original WDM file and simulation period
Update WDM       Updated      1995-10-01 to 2010-12-31  Updated WDM file and simulation period
Update WDM-Sim   Updated      1995-10-01 to 2014-05-31  Updated WDM file and simulation period

Adding the new variable "var.name" as a factor to identify what file/run and "long.per" for the period. 
```{r}
df.file <- data.frame(file=unique(df.summ00$file),var.name=factor(c("Obs","Original","Update WDM","Update WDM-Sim"),levels=c("Obs","Original","Update WDM","Update WDM-Sim")))
df.summ01 <- merge(df.summ00,df.file)
df.period <- data.frame(period=unique(df.summ00$period),long.per=factor(c("Sim-period","Summer","Winter","Storm"),levels=c("Sim-period","Summer","Winter","Storm")))
df.summ01 <- merge(df.summ01,df.period)
str(df.summ01)
```


Figure 1. Sum of flow-volume for each run by period.
```{r, echo=FALSE}
## plots of totals
p00 <- ggplot(data=df.summ01) + 
  geom_bar(aes(x=var.name,y=vol.ac.ft.sum,fill=var.name),stat="identity") +
  facet_grid(~long.per) +
  theme(axis.title.x=element_blank(), 
        axis.text.x=element_blank(), 
        axis.ticks=element_blank(),
        legend.position="bottom",
        legend.title=element_blank()) +
  ylab("Sum (ac-ft)")
plot(p00)
```


Figure 2. Annual average of flow-volume for each run by period.
```{r, echo=FALSE}
p01 <- ggplot(data=df.summ01) + 
  geom_bar(aes(x=var.name,y=vol.ac.ft.mean,fill=var.name),stat="identity") +
  facet_grid(~long.per) +
  theme(axis.title.x=element_blank(), 
        axis.text.x=element_blank(), 
        axis.ticks=element_blank(),
        legend.position="bottom",
        legend.title=element_blank()) +
  ylab("Annual Mean (ac-ft)")
plot(p01)  
```


```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="ann",c("var.name","vol.ac.ft.sum")],
      format="markdown",row.names=FALSE)
```


```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="smr",c("var.name","vol.ac.ft.sum")],
      format="markdown",row.names=FALSE)
```

```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="wtr",c("var.name","vol.ac.ft.sum")],
      format="markdown",row.names=FALSE)
```

```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="stm",c("var.name","vol.ac.ft.sum")],
      format="markdown",row.names=FALSE)
```


```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="ann",c("var.name","vol.ac.ft.mean")],
      format="markdown",row.names=FALSE)
```


```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="smr",c("var.name","vol.ac.ft.mean")],
      format="markdown",row.names=FALSE)
```

```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="wtr",c("var.name","vol.ac.ft.mean")],
      format="markdown",row.names=FALSE)
```

```{r, results='asis'}
x<-kable(df.summ01[df.summ01$period=="stm",c("var.name","vol.ac.ft.mean")],
      format="markdown",row.names=FALSE)
```
